---
title: "Practical_Machine_Learning_Project"
author: "Pramod Verma"
date: "11 July 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown for the Machine Learning Project

#Loading all the necessary library for this project.


```{r warning=FALSE}
library(caret)
library(ggplot2)
library(tidyverse)
library(randomForest)
library(rpart)
library(gbm)
library(rattle)

```


# Connecting to the data url and loading it in the data frame


```{r include=TRUE}
url_train <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_test <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

raw_train <- read.csv(url(url_train))
raw_test <- read.csv(url(url_test))
```

# Removing the na, NaN and all columns with near zero variance

```{r include=TRUE}
nearzero_var <- nearZeroVar(raw_train)
training <- raw_train[, -nearzero_var]

# Still, there are lot of missing values in many columns
# Remove all variables with more than half "na" values to reduce no of columns

training <- training[,colMeans(!is.na(training))>0.5]
training <- training[,-1] # Remove the first column which is index
dim(training)

```

# Devide the clean traing data into two sets, mytrain & mytest

```{r include=TRUE}
set.seed(2341)

inTrain <- createDataPartition(training$classe, p=0.7, list = FALSE)
newtrain <- training[inTrain, ]
newtest <- training[-inTrain,]

dim(newtrain)
dim(newtest)

```


# Make the downloaded testing data columns same as training data

```{r include=TRUE}

traincol <- colnames(training[, -58]) # Last column is different
testing <- raw_test[traincol]

classe <- as.factor(x = "") # Add last column as classe in testing
testing <- cbind(testing, classe)

dim(testing)

```


# Applying the Random Forest for Predicting Classe

```{r include=TRUE}

set.seed(1717)  # Setting the seed for reproducibility

Fit1 <- train(classe ~ ., data=newtrain, method="rf")
Pred1 <- predict(Fit1, newdata=newtest)

Result1 <- confusionMatrix(Pred1, newtest$classe)
Result1  # Display the confusion matrix

plot(Fit1)  # Plot the random forest model

```

# Applying the GBM for Predicting Classe

```{r include=TRUE}

set.seed(1718)  # Setting the seed for reproducibility

Fit2 <- train(classe ~ ., data=newtrain, method="gbm")
Pred2 <- predict(Fit2, newdata=newtest)

Result2 <- confusionMatrix(Pred2, newtest$classe)
Result2  # Display the confusion matrix

plot(Fit2)  # Plot the random forest model

```


# Applying the Decision Tree for Predicting Classe

```{r include=TRUE}

trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
set.seed(17123)  # Setting the seed for reproducibility

Fit3 <- train(classe ~ ., data=newtrain, method="rpart", trControl=trctrl, tuneLength=10)

Pred3 <- predict(Fit3, newdata=newtest)

Result3 <- confusionMatrix(Pred3, newtest$classe)
Result3  # Display the confusion matrix

# Plot the random forest model
plot(Fit3)  
plot(Fit3$finalModel)
text(Fit3$finalModel, all = TRUE)
```
`
# Applying the model on the testing data
# The best model was random forest with the higest accuracy

```{r include=TRUE}

set.seed(1730)  # Setting the seed for reproducibility

Pred_Testing <- predict(Fit1, newdata=testing)

Final_Result <- data.frame(1:20,Pred_Testing)
names(Final_Result) <- c("Problem_ID", "Classe")
Final_Result

```

# End of Project
